The \emph{TestIdeals} package for \emph{Macaulay2}

### 2019-06-03 Fsing/FThresholdsPaper/FThresholds.tex
Started the FThresholds paper.

\begin{abstract}
	This note describes a \emph{Macaulay2} package for computations in prime characteristic commutative algebra.  This includes Frobenius powers and roots, $p^{-e}$-linear and $p^{e}$-linear  maps,
  singularities defined in terms of these maps, different types of test ideals and modules, and ideals compatible with a given $p^{-e}$-linear map.
\end{abstract}

@@

This paper describes methods for computing objects and numerical invariants in prime characteristic commutative algebra, implemented in the package \emph{TestIdeals} for the computer algebra system \emph{Macaulay2} \cite{M2}.
A ring $R$ of prime characteristic $p$ comes equipped with the Frobenius endomorphism
\[ F: R \to R \ \text{ given by } F(x) =  x^p,\]
which is the basis for many constructions and methods.
Notably, the Frobenius endomorphism can be used to detect whether a ring is regular \cite{KunzCharacterizationsOfRegularLocalRings}, and further, to quantify how far a ring is from being regular, measuring the severity of a singularity.

In this direction, two notable applications of the Frobenius endomorphism are the theory of tight closure
(see \cite{HochsterHunekeTC1,HochsterFoundations} for an introduction)
and the resulting theory of test ideals
(see the survey \cite{SchwedeTuckerTestIdealSurvey}).  These methods are used by a wide group of commutative algebraists and algebraic geometers.
%That said, it remains a fundamental tool in the expanding area of prime characteristic commutative algebra.

The \emph{TestIdeals} package was started during a \emph{Macaulay2} development workshop in 2012, hosted by Wake Forest University.
The package, at that time called \emph{PosChar}, aimed at providing a unified and efficient set of tools to study singularities in characteristic $p > 0$, and in particular, collecting and implementing several algorithms that had been described in research papers.
Since then, \emph{PosChar} was split into two packages, \emph{TestIdeals} and \emph{FThresholds}, and
much more functionality has been added by many contributors, during several more \emph{Macaulay2} development workshops.\footnote{Development workshops hosted by the University of California, Berkeley (2014, 2017), Boise State University (2015), and the University of Utah (2016).}

%{\color{blue} [Maybe we don't need to list all examples?  Several other algorithms were implemented independently and incorporated into the package.]}
Starting at least with Kunz \cite{KunzCharacterizationsOfRegularLocalRings} and Fedder \cite{FedderFPureRat}, it has been known that the Frobenius endormorphism offers \emph{effective} methods for measuring singularities in positive characteristic.
However, the algorithms that form the basis of this package are the methods for computing Frobenius roots, $\star$-closures, and parameter test ideals.
These first appeared in \cite{KatzmanParameterTestIdealOfCMRings,BlickleMustataSmithDiscretenessAndRationalityOfFThresholds,BlickleMustataSmithFThresholdsOfHypersurfaces,KatzmanFrobeniusMapsOnInjectiveHulls}.
Another algorithm in \cite{KatzmanSchwedeAlgorithm} for computing prime ideals compatible with a given $p^{-e}$-linear map was implemented and used to produce the examples in that paper.  The methods for computing test ideals and test modules that were used implicitly in papers such as \cite{BlickleSchwedeTakagiZhang,KatzmanLyubeznikZhangOnDiscretenessAndRationality,SchwedeTuckerTestIdealFiniteMaps} became implementable via the Frobenius roots functionality.
Algorithms for computing $F$-pure thresholds from \cite{HernandezFInvariantsOfDiagonalHyp}, \cite{HernandezFPureThresholdOfBinomial}, and \cite{HernandezTeixeiraFThresholdFunctions} form some of the key methods in the forthcoming \emph{FThresholds} package.
%{\hfill\large\color{red} [Expand: how were the test module methods conceived?]}\\

\subsection*{Acknowledgements}
The authors thank other contributors of the \emph{TestIdeals} package, namely
Erin Bela, Juliette Bruce, Drew Ellingson, Matthew Mastroeni, and Maral Mostafazadehfard.
This paper, and the finishing touches to the \emph{TestIdeals} package, were made at the University of Utah in 2018, and the visiting authors thank the Department of Mathematics for its hospitality.  The authors also thank the referees for their numerous useful comments on both this paper and the code.

\section{Frobenius powers and Frobenius roots}\label{Section: Frobenius powers and Frobenius roots}

Let $R$ denote a commutative ring of prime characteristic $p$.

\begin{definition}
Given an ideal $I\subseteq R$ and an integer $e\geq 0$, we define the \emph{$p^e$-th Frobenius power of $I$}, denoted $I^{[p^e]}$, to be the ideal
generated by the $p^e$-th powers of all elements of $I$.
\end{definition}

It is easy to see that if $I$ is generated by $g_1, \dots, g_\ell$, then $I^{[p^e]}$ is generated by $g_1^{p^e}, \dots, g_\ell^{p^e}$.

\begin{definition}
Given an ideal $I\subseteq R$ and an integer $e\geq 0$, we define the \emph{$p^e$-th Frobenius root of $I$}, denoted $I^{[1/p^{e}]}$, to be the smallest ideal $J$ such that $I\subseteq J^{[p^e]}$, if such an ideal exists.
\end{definition}

Frobenius roots always exist in polynomial and power series rings, and in $F$-finite regular rings
(cf.~\cite[\S 2]{BlickleMustataSmithDiscretenessAndRationalityOfFThresholds} and \cite[\S 5]{KatzmanParameterTestIdealOfCMRings}).
Below is an example of a computation of a Frobenius root in a polynomial ring.
In \autoref{ss: math behind} we describe the main ideas behind such computations.

\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z];

i2 : I = ideal(x^6*y*z + x^2*y^12*z^3 + x*y*z^18);

o2 : Ideal of R

i3 : frobeniusPower(1/5, I)

                2   3
o3 = ideal (x, y , z )

o3 : Ideal of R
\end{MyVerbatim}
}
\medskip

\subsection{The mathematics behind the computation of Frobenius roots}
\label{ss: math behind}

We can also describe Frobenius roots as follows:  In a (sufficiently local) regular ring, we have an identification of $R$ with its canonical module $\omega_R$.
On the other hand, the Grothendieck dual to the $e$-iterated Frobenius map\footnote{In a reduced ring $R$, it is often convenient to express the Frobenius endomorphism $R \to R$ as the inclusion $R \hookrightarrow R^{1/p}$, so that the source and target  are distinguished.} $R \to R^{1/p^e}$ provides a map
\begin{equation}
\label{eq.DualToFrobenius}
T : \omega_{R^{1/p^e}} \to \omega_R,
\end{equation}
which using the identification $R \cong \omega_R$ gives us a map
\[
T : R^{1/p^e} \to R.
\]
It is not difficult to prove that
$T(I^{1/p^e})=I^{[1/p^e]}$,
where $I^{1/p^e} \subseteq R^{1/p^e}$ is simply the ideal of $p^e$-th roots of the elements of $I$.
In the case that
\[
R = \mathbb{K}[x_1, \ldots, x_d],
\]
where $\mathbb{K}$ is a perfect field,
$R^{1/p^e}$ is a free $R$-module with basis consisting of the monomials
\[
\mathbf{x}^{\boldsymbol{\lambda}/p^e}= x_1^{\lambda_1/p^e} \cdots x_d^{\lambda_d/p^e},
\]
where $0 \leq \lambda_i \leq p^e-1$.
Furthermore, the map $T : R^{1/p^e} \to R$ is simply the projection defined as follows:
\[
T(x_1^{\lambda_1/p^e} \cdots x_d^{\lambda_d/p^e}) = \left\{ \begin{array}{rl} 1 &\text{if } \lambda_i = p^{e}-1 \text{ for all }i \\ 0 & \text{otherwise}  \end{array} \right.
\]
Using this, it is not difficult to see that if
\begin{equation}
\label{eqn1}
f^{1/p^e} = \sum_{\boldsymbol{\lambda}} f_{\boldsymbol{\lambda}} \mathbf{x}^{\boldsymbol{\lambda} / p^e},
\end{equation}
where $\boldsymbol{\lambda}$ runs over the tuples $(\lambda_1, \dots, \lambda_d)$ with $0 \leq \lambda_j \leq p^e-1$, and $f_{\boldsymbol{\lambda}} \in R$, then
$(f)^{[1/p^e]} = T( (f)^{1/p^e})$ is the ideal generated by the $f_{\boldsymbol{\lambda}}$.
We can then compute $I^{[1/p^e]}$ for more general $I$ by linearity: if $I=(f_1,\ldots,f_m)$, then $I^{[1/p^e]}= \sum(f_i)^{[1/p^e]}$.

\subsubsection{Complexity}
The computation of Frobenius roots is the workhorse behind many of the methods in \emph{TestIdeals}.
Hence it is important to understand how this is implemented, and its computational complexity.

The computation of Frobenius roots of ideals reduces to the case of principal ideals, and its complexity grows linearly with the number of generators of the ideal.
Furthermore the calculation of  $(f)^{[1/p^e]}$ reduces to finding the terms in the right hand side of (\ref{eqn1}), which essentially amounts to
taking each term in $f$, computing the $p^e$-th root of its coefficient, and dividing the monomial exponent vector by $p^e$ with remainder.
Hence the complexity of computing $(f)^{[1/p^e]}$ is proportional to the number of terms in $f$, and is independent of its degree.
We emphasize that the calculation of Frobenius roots \emph{does not involve the calculation of Gr\"obner bases.}

\subsection{Dual to Frobenius on quotient rings and Frobenius on top local cohomology}
\label{subsec.DualToFrobeniusOnQuotientRings}
For \emph{any} reduced ring $R$ with finite $e$-iterated Frobenius map identified with $R \hookrightarrow R^{1/p^e}$, we always have a map
\[
T_R : \omega_{R^{1/p^e}} \to \omega_R.
\]
At a maximal ideal $\fram$ of height $d$, this map is Matlis dual to the Frobenius map on $H^d_{\fram}(R)$.
Hence we can study local cohomology by studying this map.
If $R = S/I$, where $S$ is a polynomial ring over a finite field, we can implement this map \cite{FedderFPureRat, KatzmanParameterTestIdealOfCMRings}.  We briefly explain the case of a hypersurface here.

If $I = (f)$, then again $\omega_R \cong R = S/I$, and the action of $T_R$ can be computed on $S$.  In fact, if $\overline{J} \subseteq R \cong \omega_R$ has preimage  $J \subseteq S$,
then setting $u = f^{p^e-1}$, we have that $T_R({\overline J}^{1/p^e})$ is the image of $(u J)^{[1/p^e]}$ in $R$.
More generally, for non-hypersurfaces, the analog to $u$ is chosen to be an element that maps onto the generator of the cyclic $R$-module
\[
\frac{(I^{[p^e]} : I) \cap (\Omega^{[p^e]} : \Omega)}{ I^{[p^e]}},
\]
where $\Omega$ is an ideal of $S$ with $I\subseteq \Omega$ and whose image in $R$ is isomorphic to the canonical module of $R$; see \cite{KatzmanParameterTestIdealOfCMRings} for details.
Here we compute two examples.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : S = ZZ/5[x,y,z];

i2 : f = x^3 + y^3 + z^3;

i3 : u = f^(5 - 1);

i4 : frobeniusRoot(1, ideal u)

o4 = ideal (z, y, x)

o4 : Ideal of S

i5 : S = ZZ/7[x,y,z];

i6 : f = x^3 + y^3 + z^3;

i7 : u = f^(7 - 1);

i8 : frobeniusRoot(1, ideal u)

o8 = ideal 1

o8 : Ideal of S
\end{MyVerbatim}
}\medskip

The above example has shown that the Frobenius map on top local cohomology of the cone over the Fermat elliptic curve is injective in characteristic $7$ (the dual map $\omega_{R^{1/p}} \to \omega_R$ is surjective) and not injective in characteristic $5$ (the dual map $\omega_{R^{1/p}} \to \omega_R$ is not surjective).

\subsection{A generalization of Frobenius powers and roots}

Let $R$ be an $F$-finite regular ring.
We can extend the definition of Frobenius powers as follows.

\begin{definition}[\cite{HernandezTeixeiraWittFrobeniusPowers}]
Let  $I\subseteq R$ be an ideal.
\begin{enumerate}
 \item[(a)] If $n$ is a positive integer with base $p$ expansion  $n=d_0 + d_1 p +  \dots + d_r p^r$, we define
\[ I^{[n]}=I^{d_0} (I^{d_1})^{[p]} \cdots  (I^{d_r})^{[p^r]}.\]
 \item[(b)] If $t$ is a nonnegative rational number of the form $t = a/p^e$, we define  $I^{[t]} = {\big(I^{[a]}\big)}^{[1/p^e]}.$
 \item[(c)] If $t$ is any nonnegative rational number, and $\{a_n/p^{e_n}\}_{n\geq 1}$ is a sequence of rational numbers converging to $t$ from above, we define $I^{[t]}$
 to be the stable value of the non-decreasing chain of ideals $\{I^{[a_n/p^{e_n}]}\}_{n\geq 1}$.
\end{enumerate}
\end{definition}

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : p = 3;

i2 : R = ZZ/p[x,y];

i3 : m = monomialIdeal(x, y);

o3 : MonomialIdeal of R

i4 : I = m^5;

o4 : MonomialIdeal of R

i5 : t = 3/5 - 1/(5*p^3);

i6 : frobeniusPower(t, I)

             2        2
o6 = ideal (y , x*y, x )

o6 : Ideal of R

i7 : frobeniusPower(t - 1/p^5, I)

o7 = ideal (y, x)

o7 : Ideal of R
\end{MyVerbatim}
}
\medskip

The generalized Frobenius powers of an ideal are test ideals (see \autoref{Section: Test Ideals}) of sufficiently general linear combinations of generators of that ideal.
Thus, the ideals computed above are test ideals of generic quintics in two variables in characteristic $p=3$, and these computations suggest that $\frac35-\frac1{5p^3}$ is a higher $F$-jumping exponent of such quintics, which is indeed the case for any $p\equiv 3 \bmod 5$ (see \cite{hernandez+etal.frobenius_examples}).
To illustrate, compare the computations above with the following.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i8 : S = ZZ/3[a..f,x,y];

i9 : G = a*x^5 + b*x^4*y + c*x^3*y^2 + d*x^2*y^3 + e*x*y^4 + f*y^5;

i10 : testIdeal(t, G)

              2        2
o10 = ideal (y , x*y, x )

o10 : Ideal of S

i11 : testIdeal(t - 1/p^5, G)

o11 = ideal (y, x)

o11 : Ideal of S
\end{MyVerbatim}
}
\medskip

\begin{caveat}
   The computations of Frobenius powers with exponents whose denominators are not powers of the characteristic $p$ are often very slow, and in several instances it is more efficient to introduce more variables and compute the test ideal of a ``very general'' linear combination of the generators of the ideal, as above.
   The code used for computations of Frobenius powers needs to undergo significant optimization.
\end{caveat}

\subsection{Frobenius  roots of submodules of free modules}

Let $S$ be a polynomial ring or power series ring.
Given a submodule $M$ of the free module $S^k$,
%
%\pedro{What are the assumptions on $S$ here?}
%
there is a smallest submodule $N$ of $S^k$ that contains $M$, for which $M\subseteq N^{[p^e]}$.
Here,  $N^{[p^e]}$ is the submodule  of $S^k$ generated by the vectors of $N$ with all coordinates raised to the $p^e$-th power. (Cf.~\cite{KatzmanZhangAlgorithm}).
Extending the previous definitions,
we call this $N$ the $p^e$-th Frobenius root of $M$, and denote it by $M^{[1/p^e]}$.

\begin{example}
Let $R=\ZZ/p\ZZ[a,b,c,d]$, and consider the ideals $\mathfrak{m}=(a,b,c,d)$ and
$$I= (a,b) \cap (a,c) \cap (c,d) \cap (c+d, a^3+b d^2).$$
Then $R/I$ is a $3$-dimensional non-Cohen--Macaulay ring.
Matlis Duality applied to $H^2_{\mathfrak{m}} (R/I)$ with its natural Frobenius map
yields a $p^{-1}$-linear map $U$ on $\Ext^2_R(R/ I, R)$.
%$\Ext^2(R/ I, R) \xrightarrow{U} \Ext^2(R/I, R)$.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/2[a,b,c,d];

i2 : I = intersect(ideal(a,b),ideal(a,c),ideal(c,d),ideal(c+d,a^3+b*d^2));

o2 : Ideal of R

i3 : f = inducedMap(R^1/I, R^1/frobenius(I));

o3 : Matrix

i4 : E2 = Ext^2(f, R^1)

o4 = {-8}  | a4+abc2+abcd a2b            |
     {-10} | a2cd3        a3cd+a3d2+bcd3 |

o4 : Matrix

i5 : target E2

o5 = cokernel {-8}  | 0     a2 b2c2 |
              {-10} | c2+d2 d4 a4d2 |

                            2
o5 : R-module, quotient of R

i6 : source E2

o6 = cokernel {-4} | 0   a  bc  |
              {-5} | c+d d2 a2d |

                            2
o6 : R-module, quotient of R
\end{MyVerbatim}
}
\medskip

The Frobenius map on $H^2_{\mathfrak{m}} (R/I)$
is injective if and only the image of
$(\Image U)^{[1/p]}$ in $\Ext^2_R(R/ I, R)$
as computed above is the whole of $R^2$.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i7 : U = matrix entries E2;

             2       2
o7 : Matrix R  <--- R

i8 : A = image matrix entries relations source E2;

i9 : frobeniusRoot(1, image U)

o9 = image {-2} | 1 0 0 |
           {0}  | 0 d a |

                             2
o9 : R-module, submodule of R
\end{MyVerbatim}
}\medskip

The calculation above shows that $R/I$ is not $F$-injective (cf. \autoref{Section: F-singularities}).
Moreover, it shows that $H^2_{\mathfrak{m}} (R/I)\cong \Ann_{E^2} A^t$,
where $E$ is the injective hull of $R_{\mathfrak{m}}/ {\mathfrak{m}} R_{\mathfrak{m}}$,
and the Frobenius map on  $H^2_{\mathfrak{m}} (R/I)$ induced from Frobenius on $R$
is given by $U^t \Theta$, where $\Theta$ is the induced Frobenius on $E$.
The submodule of nilpotent elements in  $H^2_{\mathfrak{m}} (R/I)$ is
given by $\Ann_{E^2} B^t$, where $B$ is the smallest submodule of $R^2$ containing $\Image A + \Image U$ such that $U  B \subseteq B^{[p]}$.
The method \texttt{ascendModule} can be used to calculate  $B$ (see a detailed description of the similar method \texttt{ascendIdeal} in \autoref{Section: Test Ideals}).

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i10 : B = ascendModule(1, A, U)

o10 = image | 0   a  bc  |
            | c+d d2 a2d |

                              2
o10 : R-module, submodule of R
\end{MyVerbatim}
}\medskip

\end{example}

%\section{$p^{-e}$- and $p^{e}$-linear maps}\label{Section: p-linear maps}

%\begin{definition}%{\hfill\large\color{red} [Add references]}\\
%Let $M$ be an $S$-module and $e$ a nonnegative integer.
%\begin{enumerate}
% \item[(a)] A $p^{-e}$-linear map $\phi:M \rightarrow M$ is an additive map such that
% $\phi(s^{p^e} m)= s\phi(m)$ for all $s\in S$ and $m\in M$.
% \item[(b)] A $p^{e}$-linear map $\psi:M \rightarrow M$ is an additive map such that
% $\phi(s m)= s^{p^e}\phi(m)$ for all $s\in S$ and $m\in M$.
%\end{enumerate}
%\end{definition}

%The following two examples describe  prototypical
%$p^{-e}$- and $p^{e}$-linear maps.
%\begin{example}
%For any $S$-module $M$, we can construct a new $S$-module $M^{1/p^e}$ with elements $\{  m^{1/p^e} \,|\,m\in M\}$ by defining
%$m_1^{1/p^e} +  m_2^{1/p^e} =  (m_1 +  m_2)^{1/p^e}$ for all $m_1, m_2 \in M$ and
%$s  m^{1/p^e}=  (s^{p^e} m)^{1/p^e}$ for all $m\in M$ and $s\in S$.

%Consider any $\phi\in \Hom_S(M^{1/p^e}, M)$: if we identify $M^{1/p^e}$ with $M$ we can interpret $\phi$ as a $p^{-e}$-linear map.
%\end{example}

%\begin{example}
%The $e$-th Frobenius map $f:S \rightarrow S$  raising elements to their $p^e$-th power is clearly $p^{e}$-linear.
%Furthermore, for any ideal $I\subseteq S$ and $k\geq 0$, $f$ induces a $p^{e}$-linear map $H_I^k (S) \rightarrow H_I^k (S)$.
%\end{example}

%Let $R$ be a polynomial ring with irrelevant ideal $\mathfrak{m}$ and let $g$ be a nonzero element of $R$.
%Let $E=E_{R_{\mathfrak{m}}}(R_{\mathfrak{m}}/\mathfrak{m})$ denote the injective hull of $R_{\mathfrak{m}}/\mathfrak{m}$.

%The Frobenius map on $R$ induces a Frobenius map $S$ and on
%$H^{\dim R-1}_{\mathfrak{m}} (S)=E_S(S/\mathfrak{m}S)=\Ann_E g$ and the kernel of this map is given by
%$\Ann_E (g^{p-1}R)^{[1/p]}$ (cf.~\cite[\S 5]{KatzmanParameterTestIdealOfCMRings}).

%Thus we see that the induced $p^{e}$-linear map on $H_{(x,y,z)}^{2} \left( \mathbb{K}[x,y,z]/(x^3+y^3+z^3) \right)$ is injective
%when the characteristic of $\mathbb{K}$ is $7$ and non-injective when the characteristic is $5$.

%\subsection{Implementation and complexity}

%Let $R=\mathbb{K}[x_1, \dots, x_n]$ and fix $B$ to be $\mathbb{K}$-basis for $\mathbb{K}^{1/p^e}$.
%It is not hard to see that $R^{1/p^e}$ is a free $R$-module with free basis
%$$\left\{ b x_1^{\alpha_1/p^e} \dots x_n^{\alpha_n/p^e} \,|\, b \in B, 0\leq \alpha_1, \dots, \alpha_n < p^e \right\}.$$

%\begin{proposition}(cf. \cite[\S 5]{KatzmanParameterTestIdealOfCMRings})
%\begin{enumerate}
%\item[(a)] If the ideal $I\subseteq R$ is generated by $g_1, \dots, g_k\in R$, then $I^{[1/p^e]}=\sum_{i=1}^k (g_i R)^{[1/p^e]}$.
%\item[(b)] If
%\begin{equation}\label{eqn1}
%g=\sum_{b\in B, \atop{0\leq \alpha_1, \dots, \alpha_n < p^e}} r_{b, \alpha_1, \dots, \alpha_n} \left(b x_1^{\alpha_1/p^e} \dots x_n^{\alpha_n/p^e}\right),
%\end{equation}
%then $(gR)^{[1/p^e]}$ is generated by $\{ r_{b, \alpha_1, \dots, \alpha_n} \,|\, b\in B, 0\leq \alpha_1, \dots, \alpha_n < p^e \}$.
%\end{enumerate}
%\end{proposition}

\section{$F$-singularities}\label{Section: F-singularities}

The \emph{TestIdeals} package includes methods for determining if a ring is $F$-injective, $F$-pure, $F$-rational, or $F$-regular.

\subsection{$F$-injectivity}

\begin{definition}
A local ring $(R, \mathfrak{m})$ is called \emph{F-injective} if the map
$H^{i}_{\mathfrak{m}}(R) \rightarrow H^{i}_{\mathfrak{m}}(R^{1/p})$ is
injective for all $i >0$. An arbitrary ring is called \emph{$F$-injective} if its
localization at each prime ideal of $R$ is $F$-injective.
\end{definition}

The function \texttt{isFInjective} determines whether the ring $R = S/I$ is
$F$-injective, where $S$ is a polynomial ring.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/7[x,y,z]/(x^3 + y^3 + z^3);

i2 : isFInjective R

o2 = true

i3 : R = ZZ/5[x,y,z]/(x^3 + y^3 + z^3);

i4 : isFInjective R

o4 = false
\end{MyVerbatim}
}\medskip

Equivalently, a ring is $F$-injective if the maps on the cohomology of the dualizing complex
\[
h^{-i} \omega_{R^{1/p}}^{\mydot} \to h^{-i} \omega_R^{\mydot}
\]
surject for all $i$.
Note that $h^{-i} \omega_R^{\mydot} \cong \Ext^{\dim S - i}(R, S)$, the latter of which \emph{Macaulay2} readily computes.
The algorithm used by \texttt{isFInjective} works by checking the surjectivity of the dual Frobenius map
\[
\Ext^{\dim S - i}(R^{1/p}, S) \to \Ext^{\dim S - i}(R, S).
\]
%using the fact that the Matlist dual of \Ext^{i}(\blank, S)$ \cong H^{\dim S -
%i}_{\mathfrak{m}}(\blank) $.
We begin by computing the map $R
\rightarrow R^{1/p}$ using the \emph{PushForward} package \cite{PushForward}.
Next the algorithm computes
$\Ext^{i}( \blank, S)$ applied to the map from the previous step.  Then $R$ is $F$-injective precisely when the
cokernel of $\Ext^{i}( \blank, S)$ is trivial for $i$.

The Frobenius action on top local cohomology (dual to $\omega_R$) is usually computed in a different (faster) way than the other cohomologies, and this is modified by the \texttt{CanonicalStrategy} option.
The default value for this option is \texttt{Katzman}, which instead of using the \emph{PushForward} package, relies on the fact that we already know how to compute the Frobenius action on the canonical module, as described in \autoref{subsec.DualToFrobeniusOnQuotientRings}.

The performance of the algorithm can be improved if the ring of interest is
nice enough. If the ring is Cohen--Macaulay, then setting \texttt{AssumeCM =>
true} (the default value is \texttt{false}) lets the algorithm check the Frobenius action only on top cohomology
(which is typically much faster, as explained above).
When studying a reduced ring,  setting \texttt{AssumeReduced => true} (the default value) avoids
computing the bottom local cohomology, and when studying a normal ring, setting
\texttt{AssumeNormal => true} (the default is \texttt{false}) avoids computing the bottom two local
cohomologies.

By default the algorithm checks for $F$-injectivity at all points of $\Spec R$.  However, one
can choose to check $F$-injectivity only at the origin by setting the
option \texttt{IsLocal} to  \texttt{true}.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/7[x,y,z]/((x - 1)^5 + (y + 1)^5 + z^5);

i2 : isFInjective R -- R is not globally F-injective...

o2 = false

i3 : isFInjective(R, IsLocal => true) -- but is F-injective at the origin

o3 = true
\end{MyVerbatim}
}\medskip

\subsection{$F$-regularity}

\begin{definition}[\cite{HochsterHunekeTC1,HaraWatanabeFRegFPure}]
A ring $R$ is called \emph{strongly $F$-regular} if the (big) test ideal $\tau(R)$ is equal to $R$.  Likewise a pair $(R, f^t)$ is called \emph{strongly $F$-regular} if $\tau(R, f^t) = R$.
\end{definition}

The command \texttt{isFRegular} checks whether a ring or pair is strongly
$F$-regular. Below are two examples, one of which is $F$-regular and the other one is not.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y,z]/(x^2 + y*z);

i2 : isFRegular R

o2 = true

i3 : R = ZZ/7[x,y,z]/(x^3 + y^3 + z^3);

i4 : isFRegular R

o4 = false
\end{MyVerbatim}
}\medskip

We can also check whether a pair $(R, f^t)$ is $F$-regular.

\medskip
{\small\setstretch{.67}
\begin{MyVerbatim}
i1 : R = ZZ/5[x,y];

i2 : f = y^2 - x^3;

i3 : isFRegular(1/2, f)

o3 = true

i4 : isFRegular(5/6, f)

o4 = false

i5 : isFRegular(4/5, f)

o5 = false

i6 : isFRegular(4/5 - 1/100000, f)

o6 = true
\end{MyVerbatim}
}\medskip

All of these checks are done by actually computing the test ideal, as described in \autoref{Section: Test Ideals}.

If the input ring is $\mathbb{Q}$-Gorenstein, then in each of the cases above, the output is a boolean indicating if the
ring is strongly $F$-regular. If the input ring is not
$\mathbb{Q}$-Gorenstein, then the algorithm can be used to determine if a
ring is strongly $F$-regular, but cannot prove that a ring is not strongly
$F$-regular (this latter functionality can, however, be enabled by setting \texttt{QGorensteinIndex => infinity}).

In the case that $R$ is $\mathbb{Q}$-Gorenstein, the algorithm works by
computing the test ideal $\tau$ of the ring  (or the pair) using \texttt{testIdeal}
and checking whether $\tau=R$.
In the non-$\mathbb{Q}$-Gorenstein case, the algorithm checks for
strong $F$-regularity by computing better and
better approximations of the test ideal, and checking whether any of these is the unit ideal.
To compute approximations of the
test ideal, the algorithm computes a test element $c$ with \texttt{testElement}
and then uses \texttt{frobeniusRoot} to compute the $e$-th root of
$c(I^{[p^{e}]} : I)$; appropriate modifications are made for pairs. If at
any step the approximation is the unit ideal, then then the algorithm
returns \texttt{true}. Otherwise the algorithm continues checking for each $e$
until a specified limit is reached. The default limit is 2, and can be
changed using the option \texttt{DepthOfSearch}.

A number of options can be used to speed up the performance of some of the
internal functions. The option \texttt{AssumeDomain} can be used if $R$ is an
integral domain, \texttt{FrobeniusRootStrategy} chooses a strategy for
internal \texttt{frobeniusRoot} calls, \texttt{MaxCartierIndex} sets the
maximum Gorenstein index to search for when working with a
$\mathbb{Q}$-Gorenstein ambient ring, and \texttt{QGorensteinIndex}
allows the user to specify the $\mathbb{Q}$-Gorenstein index of the ring.

>>>>

\begin{abstract}
	This note describes a \emph{Macaulay2} package for computations of $F$(-pure) thresholds in characteristic $p > 0$ commutative algebra.  
\end{abstract}

@@

This paper describes a Macaulay2 package which provides tools to compute, $F$-pure thresholds, $F$-thresholds and $F$-jumping numbers.  We begin with some history

{\it Acknowledgements}  The authors thank...

\section{The $F$-pure threshold limit and {\tt nu}}

Suppose $R = k[x_1, \dots, x_n]$ is a polynomial ring over a finite field and $f \in R$.    If $f^a \notin \langle x_1^{p^e}, \dots, x_n^{p^e} \rangle = \fram^{[p^e]}$ then the pair $(R, f^{a/(p^e -1)})$ is $F$-pure at the origin, whereas if $f^a \in \fram^{[p^e]}$, then the pair $(R, f^{a/p^e})$ is not $F$-pure at the origin.  By finding the $a \in \bZ$ where that transition happens, and then limiting over $e$, one can compute the $F$-pure threshold.  See \cite{MustataTakagiWatanabeFThresholdsAndBernsteinSato}.

We set $\nu(e, f)$ to be largest $a \in \bZ$ such that $f^a \notin \fram^{[p^e]}$ (note in the literature, it is frequently denoted by $\nu_f(p^e)$.  It immediately follows that 
\[
\fpt(f) = \lim_{e \to \infty} {\nu(e, f) \over p^e}.
\]
These numbers $\nu(e, f)$ can be computed using the command {\tt nu}, as in the example below.  
\medskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i1 : loadPackage "FThresholds"

o1 = FThresholds

o1 : Package

i2 : R = ZZ/7[x,y]

o2 = R

o2 : PolynomialRing

i3 : f = y^2 - x^3

        3    2
o3 = - x  + y

o3 : R

i4 : nu(1, f)

o4 = 5

o4 : QQ

i5 : nu(2, f)

o5 = 40

o5 : QQ

i6 : nu(3, f)

o6 = 285

o6 : QQ
\end{MyVerbatim}
}
\medskip
The numbers $5/7, 40/49, 285/343$ then approximate the $F$-pure threshold, which in this case happens to equal $5/6$.  By default, if the function $f$ has a special form (is diagonal or binomial), then the function does not check whether $f^a$ is in  $\langle x_1^{p^e}, \dots, x_n^{p^e} \rangle$, but instead uses general formulas for $\nu$s, see \cite{HernandezFPureThresholdOfBinomial,HernandezFInvariantsOfDiagonalHyp}.

In fact, one can make the following more general definition.  If $I$ and $J$ are two ideals (with $I \subseteq \sqrt{J}$), one can define
\[
\nu(e, I, J) = \max \{ n \in \bZ \; |\; I^n \not\subseteq J^{[p^e]} \}.
\]
This is frequently denoted by $\nu_I^J(p^e)$ in the literature.  In the case that $I = \langle f \rangle$ and $J = \fram$, this just recovers the $\nu$'s defined above; the limit of which is the $F$-pure threshold.  For more general $I$ and $J$, the limit
\[
\lim_{e \to \infty} {\nu(e, I, J) \over p^e}
\]
is called the \emph{$F$-threshold of $I$ with respect to $J$}.  To compute these values we use a binary search, unless the user specifies that we should use a linear search instead {\tt Search=>Linear}.

There is one more option that deserves mentioning, {\tt ContainmentTest}.  There are two ways to check whether $f^a$ (or $I^a$) is contained in $J^{[p^e]}$.  In general it is true that
\[
I^a \subseteq J^{[p^e]} \text{ $\Leftrightarrow$ } (I^a)^{[1/p^e]} \subseteq J
\]
where $\bullet^{[1/p^e]}$ is defined as in the {\tt TestIdeals} package.  These are controlled by {\tt ContainmentTest => StandardPower} and {\tt ContainmentTest => FrobeniusRoot} respectively.
For principal $I$, the command on the right is generally faster (and so {\tt FrobeniusRoot} is the default strategy for computing $\nu$).  For non-principal $I$, frequently the command on the left is quicker, and so {\tt StandardPower} is default strategy.

Finally, the option {\tt ContainmentTest => FrobeniusPower}, then ${\tt \nu(e, I, A)}$ computes the largest integer $n$ such that
\[
I^{[p^n]} \subseteq J^{[p^e]}.
\]
See \cite{HernandezTeixeiraWittFrobeniusPowers} for more discussion of this.

\section{{\tt isFPT} and {\tt isFJumpingExponent}}

The command {\tt isFPT(t, f)} checks whether $t$ is the $F$-pure threshold of $f$.  In general this does this by computing two test ideals:
\[
\tau(R, f^t) \text{ and } \tau(R, f^{t-\epsilon}).
\]
It follows that $\fpt(f) = t$ if and only if $\tau(R, f^t) \neq R$ and $\tau(R, f^{t-\epsilon}) = R$.  The function computes these two values by using the {\tt testIdeal} and {\tt FPureModule} commands from the {\tt TestIdeals} package.  Note that in a strongly $F$-regular Gorenstein ambient ring, {\tt FPureModule(t, f)} computes exactly $\tau(R, f^{t-\epsilon})$.  These commands work more generally, but they \emph{only} currently work when $R$ is $\bQ$-Gorenstein with index not divisible by the characteristic $p > 0$.

The command {\tt isFJumpingExponent} works similarly, a value $t$ is an $F$-jumping exponent if and only if
\[
\tau(R, f^t) \neq \tau(R, f^{t-\epsilon}).  
\]
Again, this command works as long as $R$ is $\bQ$-Gorenstein with index not divisible by $p$.

\section{\tt fpt}

The core function in this package is the {\tt fpt} function.

\section{}
